---
title: "R Day 2"
author: "Harish Chockalingam"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
here::here("data","gapminder5.csv")

gapminder<- read.csv(here::here("data", "gapminder5.csv"))
```

## Converting factors to characters

```{r}
gapminder$country<-as.character(gapminder$country)
gapminder$continent<-as.character(gapminder$continent)
str(gapminder)
```

obs<-nrow(gapminder)
for (i in obs) {
  gapminder[i, "gdp"]<-gapminder[i,"pop"]*gapminder[i,"gdpPercap"]
}

## Creating log variable with for loop
```{r}
obs<-nrow(gapminder)
for(i in obs){
 gapminder[i,"log_gdpPerCap"]<-log(gapminder[i,"gdpPercap"])
 gapminder[i,"log_pop"]<-log(gapminder[i,"pop"])
}

years<-unique(gapminder$year)
for(i in years){
  mean_le<-mean(gapminder$lifeExp[gapminder$year==i],na.rm=T)
  print(paste0(i,": ",mean_le))
}

conts <- unique(gapminder$continent)
for (i in conts){
  mean_le<-mean(gapminder$lifeExp[gapminder$continent==i], na.rm=T)
  print((i,": ", mean_le))
}

for(i in conts){
  print(paste0("Continent: ",i))
  for(j in years){
    mean_le<-mean(gapminder$lifeExp[gapminder$continent==i & mean_legapminder$year==j], na.rm=T)
    print(paste0(j, ": ", mean_le))
  }
}

for(i in conts){
  print(paste0("Continent: ", i))
  for(j in years){
    sd_le<-sd(gapminder$lifeExp[gapminder$continent==i & gapminder$year==j],na.rm=T)
    print(paste0(j,": ", sd_le))
  }
}

vars<-gapminder[,c("lifeExp","pop","gdpPercap")]
apply(vars, 2, mean)
lapply(gapminder, mean)
sapply(years,function(x) mean(gapminder$lifeExp[gapminder$year==x]))

i<-1987
while(i<2003){
  sd_lf<-sd(gapminder$lifeExp[gapminder$year==i])
print(paste0(i,": ", sd_lf))
i<-i+5
}

```
for(i in years){
  if(i>=1987){
    mean_pop<-mean(gapminder$lifeExp[gapminder$year==i])
    print(paste0(i,": ", mean_pop))
  } else {
      print("Sorry, year is less than 1987")
      }
}

reports_m4<-
function(df, variable, continent){
  var<-df[[variable]][df$continent==continent]
  max_le<-max(var)
  min_le<-min(var)
  cat("Continent:",continent,
  "/nMinimum Life expectancy:", min_le,
  "/nMaximum Life expectancy:", max_le)
}
report_stats(gapminder,"lifeExp", "Asia"")